Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2015-03-10 paper/poly.tex
complete the shifts section

Let $V$ be a set of variabls.  An index $I \in \ind(V)$ denotes a
\emph{base polynomial} $P_I : \State \to \Q$ that maps a program
state $\state$ to a rational number.  The definition of $P_I$ is
$
P_I(\state) = \prod_{(x,y)} |[\state(x),\state(y)]|^{i_{(x,y)}}.
$
Remark that following the standard convention for $\prod$, the
empty index represents the constant base polynomial 1.

@@

The previous potential annotations will be used to decorate
programs and bound their resource consumption.  A program
$P$ consumes $n$ resource units turning a state $\state$ into $\state'$,
we note this $\state \rightarrow^n_P \state'$.  The \emph{potential
method} make us look for two annotations $Q$ and $Q'$ such
that
$$
  \forall n\, \state\, \state'.\, \state \rightarrow^n_P \state'
    \implies P_Q(\state) \ge n + P_{Q'}(\state').
$$
The composition of these annotations through the whole program
using a quantitative logic permits to derive a resource bound.
In the sequel we explicit a relation that is sufficient to
satisfy the previous condition when the program $P$ is an
increment operation $x \gets x+y$.

@@

By symmetry we can consider the case where $P=x \gets x+y$ with
$y>0$, the case where $y \le 0$ is only a few signs away.  Let
$Q$ be a potential annotation, we will describe a $Q'$ and a
$P^y_.$ map such that if $\state \rightarrow_P \state'$ (with no resource cost),
$P_Q(\state) = P^y_{Q'}(\state')$.  We call $Q'$ a shift of $Q$ and
write $Q' = \qsh_x^+ Q$.  (When $y\le0$ we get $\qsh_x^-Q$.)

@@

>>>>

Let $V$ be a set of variabls.  An index $I \in \ind(V)$ denotes a
\emph{base polynomial} $P_I : \State \to \Q$ that maps a program
state $\state$ to a rational number.  The definition of $P_I$ is
$
P_I(\state) = \prod_{(x,y)} |[\state(x),\state(y)]|^{i_{(x,y)}}.
$
Remark that following the standard convention for $\prod$, the
empty index represents the constant base polynomial 1.  We will
later use the notation $P_{(x,y)}(\state) = |[\state(x),\state(y)]|$
to define $P_I$ as $P_I = \prod_{(x,y)} P_{(x,y)}$.

@@

The previous potential annotations will be used to decorate
programs and bound their resource consumption.  A program
$C$ consumes $n$ resource units turning a state $\state$ into $\state'$,
we note this $\state \rightarrow^n_C \state'$.  The \emph{potential
method} make us look for two annotations $Q$ and $Q'$ such
that
$$
  \forall n\, \state\, \state'.\, \state \rightarrow^n_C \state'
    \implies P_Q(\state) \ge n + P_{Q'}(\state').
$$
The composition of these annotations through the whole program
using a quantitative logic permits to derive a resource bound.
In the sequel we explicit a relation that is sufficient to
satisfy the previous condition when the program $C$ is an
increment operation $x \gets x+y$.

@@

By symmetry we can consider the case where $C=x \gets x+y$ with
$y>0$, the case where $y \le 0$ is only a few signs away.  Let
$Q$ be a potential annotation, we will describe a $Q'$ and a
$P^y_.$ map such that if $\state \rightarrow_C \state'$ (with no resource cost),
$P_Q(\state) = P^y_{Q'}(\state')$.  We call $Q'$ a shift of $Q$ and
write $Q' = \qsh_x^+ Q$.  (When $y\le0$ we get $\qsh_x^-Q$.)

@@

We now define an interpretation $P_.^y$ of indices containing
$\square$:
\begin{align*}
P^y_{(\square,z)}(\state) &=
  \left| [\sigma(x){+}\sigma(y), \sigma(z)] \,\triangle\,
         [\sigma(x), \sigma(z)] \right| \\
P^y_{(z,\square)}(\state) &=
  \left| [\sigma(z), \sigma(x){+}\sigma(y)] \,\triangle\,
         [\sigma(z), \sigma(x)] \right| \\
P^y_{(\square, \square)}(\state) &= 0 \\
P^y_{(u,v)}(\state) &= |[\sigma(u), \sigma(v)]|,
\end{align*}
we use $A \,\triangle\, B$ to denote the symmetric difference
of $A$ and $B$.
The quantity $P^y_{(\square,z)}(\state)$ is the size change of
the interval $[x, z]$.  If $\state$ is the state before the
assignment and $\state'$ is the state after the assignment, we
have
$P_{(x,z)}(\state) = P^y_{(x,z)}(\state') + P^y_{(\square,z)}(\state')$
and
$P_{(z,x)}(\state) = P^y_{(z,x)}(\state') - P^y_{(z,\square)}(\state')$.
With these definitions, we have the following lemma.
\begin{lemma}
If $\state \rightarrow_C \state'$ and $\state(y)>0$ then
$P_Q(\state) = P^y_{\qsh^+_x Q}(\state')$.
\end{lemma}
